Rising Trends in Shelter Resident Deaths in Toronto Over the Past 17 Years

### 2024-09-25 paper/paper.qmd
styled used for formatting
styled used for formatting

# Showing the summary of cleaned data.
kable(summary(analysis_data[, c("year_month", "total_decedents")]),
      col.names = c("Time", "Number of decedents"))
```

@@

# Plotting a histogram for total_decedents
ggplot(analysis_data, aes(x = total_decedents)) +
  geom_histogram(binwidth = 1, fill = "steelblue", color = "black", alpha = 0.7) +
  labs(title = "Distribution of monthly number of decedents", x = "Number of decedents", y = "Frequency") +
  theme_minimal()+
  theme(
    panel.grid.major = element_line(color = "gray80"),  
    panel.grid.minor = element_blank(),  
    plot.title = element_text(hjust = 0.5, size = 12),  
    axis.title.x = element_text(size = 12), 
    axis.title.y = element_text(size = 12),
    axis.text.x = element_text(hjust = 0.5,  size = 10),
    axis.text.y = element_text(hjust = 0,  size = 10),
  )
```

@@

# plot an graph for the trend of number of decedents over time.
ggplot(analysis_data, aes(x = as.Date(paste(year_month, "-01", sep="")), y = total_decedents)) +
  geom_line(color = "steelblue", size = 0.7) +  
  labs(title = "Number of decedents over time", 
       x = "Time", 
       y = "Number of decedents") +
  scale_x_date(date_labels = "%Y", date_breaks = "2 years") +  
  theme_minimal(base_size = 15) +  
  theme(
    panel.grid.major = element_line(color = "gray80"),  
    panel.grid.minor = element_blank(),  
    plot.title = element_text(hjust = 0.5, size = 12),  
    axis.title.x = element_text(size = 12), 
    axis.title.y = element_text(size = 12),
    axis.text.x = element_text(angle = 30, hjust = 0.5,  size = 10),
    axis.text.y = element_text(hjust = 0,  size = 10),
  )
```
## Discussion of measurement

@@

# Scatter plot and fitted line
ggplot(analysis_data, aes(x = ym(year_month), y = total_decedents)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Simple linear regression of number of decedent by time",
       x = "Time",
       y = "Number of Decedents") +
  theme(
    panel.grid.major = element_line(color = "gray80"),  
    panel.grid.minor = element_blank(),  
    plot.title = element_text(hjust = 0.5, size = 12),  
    axis.title.x = element_text(size = 12), 
    axis.title.y = element_text(size = 12),
    axis.text.x = element_text(hjust = 0.5,  size = 10),
    axis.text.y = element_text(hjust = 0,  size = 10),
  )
```

@@

# Create a bar chart with percentage labels
ggplot(filtered_data, aes(x = year, y = total_decedents)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  geom_text(aes(label = paste0(round(total_decedents_pct, 1), "%")), 
            vjust = -0.5, size = 2.3) +  # Add percentage labels
  labs(title = "Number of decedents per year", x = "Year", y = "Number of decedents") +
  theme_minimal()+
  theme(
    panel.grid.major = element_line(color = "gray80"),  
    panel.grid.minor = element_blank(),  
    plot.title = element_text(hjust = 0.5, size = 12),  
    axis.title.x = element_text(size = 12), 
    axis.title.y = element_text(size = 12),
    axis.text.x = element_text(angle = 30, hjust = 0.5,  size = 10),
    axis.text.y = element_text(hjust = 0,  size = 10),

)
```

@@

# Grouping data by season
seasonal_data <- analysis_data %>%
  group_by(season) %>%
  summarize(mean_decedents = mean(total_decedents),
            total_decedents = sum(total_decedents))

@@

# Create a bar chart with percentage labels
ggplot(seasonal_data, aes(x = season, y = mean_decedents, fill = season)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = paste0(round(total_decedents_pct, 1), "%")), 
            vjust = -0.5, size = 2.5) +  # Add percentage labels
  labs(title = "Average number of decedents per season", 
       x = "Season", 
       y = "Average number of decedents") +
  theme_minimal() +
  theme(
    panel.grid.major = element_line(color = "gray80"),  
    panel.grid.minor = element_blank(),  
    plot.title = element_text(hjust = 0.5, size = 12),  
    axis.title.x = element_text(size = 12), 
    axis.title.y = element_text(size = 12),
    axis.text.x = element_text(hjust = 0.5,  size = 10),
    axis.text.y = element_text(hjust = 0,  size = 10),
  )
```

>>>>

# Showing the summary of cleaned data.
kable(summary(analysis_data[, c("year_month", "total_decedents")]),
  col.names = c("Time", "Number of decedents")
)
```

@@

# Plotting a histogram for total_decedents
ggplot(analysis_data, aes(x = total_decedents)) +
  geom_histogram(binwidth = 1, fill = "steelblue", color = "black", alpha = 0.7) +
  labs(title = "Distribution of monthly number of decedents", x = "Number of decedents", y = "Frequency") +
  theme_minimal() +
  theme(
    panel.grid.major = element_line(color = "gray80"),
    panel.grid.minor = element_blank(),
    plot.title = element_text(hjust = 0.5, size = 12),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text.x = element_text(hjust = 0.5, size = 10),
    axis.text.y = element_text(hjust = 0, size = 10),
  )
```

@@

# plot an graph for the trend of number of decedents over time.
ggplot(analysis_data, aes(x = as.Date(paste(year_month, "-01", sep = "")), y = total_decedents)) +
  geom_line(color = "steelblue", size = 0.7) +
  labs(
    title = "Number of decedents over time",
    x = "Time",
    y = "Number of decedents"
  ) +
  scale_x_date(date_labels = "%Y", date_breaks = "2 years") +
  theme_minimal(base_size = 15) +
  theme(
    panel.grid.major = element_line(color = "gray80"),
    panel.grid.minor = element_blank(),
    plot.title = element_text(hjust = 0.5, size = 12),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text.x = element_text(angle = 30, hjust = 0.5, size = 10),
    axis.text.y = element_text(hjust = 0, size = 10),
  )
```
## Discussion of measurement

@@

# Scatter plot and fitted line
ggplot(analysis_data, aes(x = ym(year_month), y = total_decedents)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Simple linear regression of number of decedent by time",
    x = "Time",
    y = "Number of Decedents"
  ) +
  theme(
    panel.grid.major = element_line(color = "gray80"),
    panel.grid.minor = element_blank(),
    plot.title = element_text(hjust = 0.5, size = 12),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text.x = element_text(hjust = 0.5, size = 10),
    axis.text.y = element_text(hjust = 0, size = 10),
  )
```

@@

# Create a bar chart with percentage labels
ggplot(filtered_data, aes(x = year, y = total_decedents)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  geom_text(aes(label = paste0(round(total_decedents_pct, 1), "%")),
    vjust = -0.5, size = 2.3
  ) + # Add percentage labels
  labs(title = "Number of decedents per year", x = "Year", y = "Number of decedents") +
  theme_minimal() +
  theme(
    panel.grid.major = element_line(color = "gray80"),
    panel.grid.minor = element_blank(),
    plot.title = element_text(hjust = 0.5, size = 12),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text.x = element_text(angle = 30, hjust = 0.5, size = 10),
    axis.text.y = element_text(hjust = 0, size = 10),
  )
```

@@

# Grouping data by season
seasonal_data <- analysis_data %>%
  group_by(season) %>%
  summarize(
    mean_decedents = mean(total_decedents),
    total_decedents = sum(total_decedents)
  )

@@

# Create a bar chart with percentage labels
ggplot(seasonal_data, aes(x = season, y = mean_decedents, fill = season)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = paste0(round(total_decedents_pct, 1), "%")),
    vjust = -0.5, size = 2.5
  ) + # Add percentage labels
  labs(
    title = "Average number of decedents per season",
    x = "Season",
    y = "Average number of decedents"
  ) +
  theme_minimal() +
  theme(
    panel.grid.major = element_line(color = "gray80"),
    panel.grid.minor = element_blank(),
    plot.title = element_text(hjust = 0.5, size = 12),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text.x = element_text(hjust = 0.5, size = 10),
    axis.text.y = element_text(hjust = 0, size = 10),
  )
```
